The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-12 manuscript/manuscript.Rnw
set grad.def in plots; include more hormone genes

@@

\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
print(ggplot(core.placenta.transcriptome.long[!grepl("(min|median|max)_expression",Species),],
             aes(y=human_name,x=capfirst(Species)))
      + geom_tile(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = "grey50",
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(), 
              axis.text.x = element_text(angle = 340, hjust = 0.1, vjust=1,face="italic"))
      )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes expressed (\ac{FPKM}$ \ge 10$) in the placenta of all organisms
  studied with housekeeping
  genes \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  removed.}
\label{fig:core_placenta_transcriptome}
\end{figure}

@@

### hormone genes
hormone.genes <- c("CGA","CGB","CGB1","CGB2","CGB5","CGB8","CSH1","CSH2","CSHL1","CYP19A1",
                   "GH2","GHRH","HSD3B1","NPPB","NPPC","PRL","RLN1")
hormone.expression.nozero <-
    gene.expression.nozero(hormone.genes)

@@

common.plot.options <- 
    list(geom_tile(aes(fill = fpkm), colour = "white"),
##         scale_fill_gradientn(colours = colorends, values = rescale(gradientends)),
         scale_fill_gradient(low = "#132B43", high = "#56B1F7", space = "Lab", na.value = "grey50",
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         scale_x_discrete("", expand = c(0, 0)),
         scale_y_discrete("", expand = c(0, 0)),
         theme_grey(base_size = 9),
         theme(legend.position = "top",
               axis.ticks = element_blank(), 
               axis.text.x = element_text(angle = 300, hjust = 0, vjust=1,face="italic")),
         facet_wrap(~type,scale="free_y",ncol=1)
         )
print(ggplot(combined.long, aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\caption{Expression of Hormone Genes, Galectins, and IGF related genes}
\label{fig:hormone_genes}
\end{figure}

>>>>

<<plot.defaults,echo=FALSE,results="hide">>=
grad.def <- list(low = "white", high = "green", space = "Lab", na.value = "gray50")

@@

\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,out.height="0.8\\textheight",echo=FALSE>>=
print(ggplot(core.placenta.transcriptome.long[!grepl("(min|median|max)_expression",Species),],
             aes(y=human_name,x=capfirst(Species)))
      + geom_tile(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low = grad.def[["low"]],
                            high = grad.def[["high"]],
                            space = grad.def[["space"]],
                            na.value = grad.def[["na.value"]],
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(), 
              axis.text.x = element_text(angle = 340, hjust = 0.1, vjust=1,face="italic"))
      )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes expressed (\ac{FPKM}$ \ge 10$) in the placenta of all organisms
  studied with housekeeping
  genes \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  removed.}
\label{fig:core_placenta_transcriptome}
\end{figure}

@@

### hormone genes
hormone.genes <- c("CGA","CGB","CGB1","CGB2","CGB5","CGB8","CSH1","CSH2","CSHL1","CYP19A1",
                   "GH2","GHRH","HSD3B1","HSD3B2","HSD17B2","HSD17B1","NPPB","NPPC","PRL","RLN1")
hormone.expression.nozero <-
    gene.expression.nozero(hormone.genes)

@@

common.plot.options <- 
    list(geom_tile(aes(fill = fpkm), colour = "white"),
##         scale_fill_gradientn(colours = colorends, values = rescale(gradientends)),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         scale_x_discrete("", expand = c(0, 0)),
         scale_y_discrete("", expand = c(0, 0)),
         theme_grey(base_size = 9),
         theme(legend.position = "top",
               axis.ticks = element_blank(), 
               axis.text.x = element_text(angle = 300, hjust = 0, vjust=1,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long, aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\caption{Expression of Hormone Genes, Galectins, and IGF related genes}
\label{fig:hormone_genes}
\end{figure}
